scoring

### 2022-04-14 darts/scoring.Rmd
add some comments

scores <- function(observed, dat) {
  logsscore <- scoringRules::logs_sample(observed, dat)
  crpsscore <- scoringRules::crps_sample(observed, dat)
  data.frame(logs = mean(logsscore[-1]), crps =  mean(crpsscore[-1]))
}
```

@@

dat <- fcst |>
  select(t = time, i = ensemble, N = value) |> 
 pivot_wider(id_cols = "t", names_from="i", values_from = "N") |> 
  select(-t) |> as.matrix()

test <- true  |>
  select(t = time, i = ensemble, N = value)

rep_scores <- 
  test |> 
  group_by(i) |> 
  group_map(~ scores(.x$N, dat)) |> 
  bind_rows()

bind_cols(rep_scores, groups)
}

rep_scores <- data |> 
  group_by(iter, ml_model, tp_model) |>
  group_map(~ score_all(.x, .y) ,.keep = TRUE) |> 
  bind_rows()
```

```{r}
# crude demo
ggplot(rep_scores) + geom_density(aes(crps, color=iter, group=iter)) + facet_grid(ml_model~tp_model)

>>>>

# trivial wrapper to score both logs and crps
scores <- function(observed, predicted) {
  logsscore <- scoringRules::logs_sample(observed, predicted)
  crpsscore <- scoringRules::crps_sample(observed, predicted)
  # drops the first point (initial condition)
  data.frame(logs = mean(logsscore[-1]), crps =  mean(crpsscore[-1]))
}
```

@@

# key trick is to make forecast a t_max x ensembles matrix for scoringRules:
predicted <- fcst |>
  select(t = time, i = ensemble, value) |> 
  pivot_wider(id_cols = "t", names_from="i", values_from = "value") |> 
  select(-t) |> 
  as.matrix()

# just select and rename relevant columns
test <- true |>  select(t = time, i = ensemble, value)

# Here we go! score each test timeseries against the entire predicted ensemble!
rep_scores <- 
  test |> 
  group_by(i) |> 
  group_map(~ scores(.x$value, predicted)) |> 
  bind_rows()

# add any "grouping variables" on as additional columns for bookkeeping
bind_cols(rep_scores, groups)
}
```

```{r}
# now actually run the scoring over all iterations, ml_models, tp_models:

bench::bench_time({
rep_scores <- data |> 
  group_by(iter, ml_model, tp_model) |>
  group_map(~ score_all(.x, .y) ,.keep = TRUE) |> 
  bind_rows()

})
```

# rough plot, probably not the best visual of this.
ggplot(rep_scores) + 
  geom_density(aes(crps, color=iter, group=iter)) + 
  facet_grid(ml_model~tp_model)
